\sysfull

### 2018-04-16 paper/relatedwork.tex
edits to related work

\textbf{Intermittently-Powered Devices.} There is a large body of research on energy harvesting and wirelessly powered embedded devices. Several recent reviews provide a broad in-depth overview~\cite{prasad_comst_2014,sample_procieee_2013,huang:commag:2015,visser_procieee_2013,kamalinejad_commag_2015,ku_cst_2016}. As computing power consumption continues to decrease, a new wave of {\em ambiently} powered systems is emerging, powered by e.g. radio waves~\cite{patel_pervasive_2017,rf_powered_computing_gollakota_2014}.
Radio-powered devices~\cite{wisp5,moo,zhao_rfid_2015,holleman_biocas_2008,thomas_jbcs_2012,naderiparizi_rfid_2015,rodriguez_tbcs_2015,liu_sigcomm_2013,kicksat,nadeau_naturebio_2017}
are related to \sys---they are the technological foundation of an
important category of intermittent computing devices. The emergence of
intermittent computing hardware platforms has led to the development of tooling
and instrumentation for such systems building~\cite{hester_sensys_2014,hester_sensys_2015,edb,stork,wisent}.

\textbf{Intermittent Execution Environments.} Early work on energy-harvesting runtime systems assumed simple, small computations not exceeding a predictable burst of energy~\cite{dewdrop} or long power-cycles/priority-based scheduling~\cite{sorber_sensys_2007}. These efforts did not consider data consistency with intermittence. Works supporting intermittent computing using dynamic, on-demand checkpoints include Mementos~\cite{mementos}, Quickrecall~\cite{quickrecall}, Hibernus++~\cite{hibernusplusplus} and HarvOS~\cite{mottola2017harvos}. Dynamic checkpointing systems may checkpoint
at any point, making it difficult to implement application-level atomicity
requirements. Moreover, some require hardware to measure stored energy, incurring energy cost. Moreover, these systems do not checkpoint non-volatile state, compromising memory consistency. DINO~\cite{dino} selectively versions non-volatile memory and checkpoints volatile state, ensuring progress and memory consistency. Rachet~\cite{ratchet} assumes all memory is non-volatile and checkpoints volatile state. Clang~\cite{hicks_isca_2017} takes a Ratchet-like approach leveraging special hardware, eliminating checkpointing and versioning overhead.

Recent task-based systems are Chain~\cite{chain} and Alpaca~\cite{alpaca}. Using static tasks, they eliminate checkpointing volatile state. Using channel-based memory models~\cite{chain} or automatic privatization and redo-logging~\cite{alpaca} they avoid checkpointing overheads. Moreover, task-based models facilitate specifying application-level atomicity properties.

\sys relates to the above efforts but is distinct in approach and mechanism. \sys relates closely to task-based models because \sys also relies on statically defined tasks to avoid checkpointing volatile state. \sys's automatic compilation spares the programmer's work of writing task-based code that prior systems demand. \sys's memory virtualization provides an alternative for ensuring memory consistency, optimized for bulk accesses to task-shared data with high locality. \sys's task coalescing optimization ameliorates the key overheads associated with prior systems.

@@

Another \sys-related prior effort is on unbounded, page-based transactional memory and deterministic parallel runtime systems~\cite{pagebasedtm,grace}. These works have a different mechanism and purpose than \sys---ensuring that data are consistent and deterministically updated during concurrent executions. What makes these
efforts related is that they manage state to ensure consistency at the
granularity of pages to amortize checking and tracking costs. \sys also tracks
accesses at the granularity of a page and uses its paging mechanism to ensure
consistency. Moreover, \sys's paging implementation, which keeps a shadow page
for each page to use during commit, is similar to the shadow paging scheme
use for transactional commit~\cite{pagebasedtm}.

>>>>

\textbf{Intermittently-powered Devices.} There is a large body of research on
energy-harvesting and wirelessly-powered embedded devices covered in recent
surveys~\cite{prasad_comst_2014,sample_procieee_2013,huang:commag:2015,visser_procieee_2013,kamalinejad_commag_2015,ku_cst_2016}.
As computing power consumption continues to decrease, a new wave of
systems that are powered by radio waves (RF) is
emerging~\cite{patel_pervasive_2017,rf_powered_computing_gollakota_2014,wisp5,moo,zhao_rfid_2015,holleman_biocas_2008,thomas_jbcs_2012,naderiparizi_rfid_2015,rodriguez_tbcs_2015,liu_sigcomm_2013,kicksat,nadeau_naturebio_2017}.
RF-powered devices are an important category of intermittent computing devices
that \sys targets.  The emergence of intermittent computing hardware platforms
has led to the development of instrumentation, debugging, and prototyping tools
for such systems
~\cite{hester_sensys_2014,hester_sensys_2015,edb,capybara,stork,wisent,flicker}.

\textbf{Checkpointing-based Systems.} Early work on energy-harvesting runtime
systems, such as Dewdrop~\cite{dewdrop}, assumed simple computations
that complete on a predictable burst of energy.
Eon~\cite{sorber_sensys_2007} has scheduled such computation based on priority.
Support for computation that spans power failures was first achieved with
statically-placed conditional checkpoints in Mementos~\cite{mementos}.

Mementos left the possibility of non-volatile state becoming inconsistent with
volatile state saved in the checkpoint, rendering in-place writes to data
structures in non-volatile memory unsafe~\cite{mspcdino}.
%
DINO~\cite{dino} addressed the consistency problem by selectively versioning
non-volatile state within the checkpoints. Ratchet~\cite{ratchet} ensured
consistency by placing a checkpoint at the beginning of each idempotent region
in the code.  Clank~\cite{hicks_isca_2017} ensured consistency with custom
hardware that dynamically tracks read-after-write dependencies in memory
accesses and checkpoints on demand. The consistency problem was also
approached with a combination of undo and redo logging in
software~\cite{baghsorkhi_cgo_2018}.
%
Just-in-time checkpointing, such as Quickrecall~\cite{quickrecall} and
Hibernus++~\cite{hibernusplusplus}, eschews inconsistency by saving all
volatile state immediately before power failure and halting the execution.
Unlike \sys, such systems rely on introspection hardware to monitor supply
voltage and on accurate worst-case bounds on checkpoint cost.

In all of the above systems, with the exception of DINO~\cite{dino},
checkpoints are dynamic, i.e. the programmer does not have explicit control
over the point at which the code may be resumed after power failure.
%
Dynamic checkpointing systems make it difficult for the programmer to respect
application-level atomicity constraints, such as correlating sensor readings.
%
Checkpointing systems that copy most volatile state scale poorly as the
size of volatile memory increases.  QuickRecall, Clank and Ratchet reduce
the copying overhead by allocating the stack in non-volatile memory, which
requires more time and energy to access than volatile memory (cf.
Section~\ref{sec:cost_task-based}) and is not a viable option for off-chip non-volatile
memory. In addition, unlike \sys, QuickRecall and Clank require custom hardware.

\textbf{Task-based Systems.} An alternative to checkpointing are recent systems
based on static tasks, such as Chain~\cite{chain}, Alpaca~\cite{alpaca} and
Mayfly~\cite{hester_sensys_2017}. Using static tasks, they eliminate
the need to checkpoint volatile state. Using channel-based memory
models~\cite{chain,hester_sensys_2017} or automatic privatization and
redo-logging~\cite{alpaca} they avoid checkpointing overheads. Moreover,
task-based models facilitate respecting application-level atomicity
constraints.
%
\sys relates closely to task-based models because \sys also relies on
statically defined tasks to avoid checkpointing volatile state. 
%
However, unlike prior systems, \sys coalesces its statically-defined tasks at
runtime into more efficient dynamic tasks that adapt to changing energy
conditions.
%
\sys's mechanism for ensuring memory consistency also differs from the
channel-based~\cite{chain} and privatization-based~\cite{alpaca} mechanisms in
prior systems. \sys keeps memory consistent through memory virtualization
optimized for bulk accesses to task-shared data with high locality.

\textbf{Task Decomposition.} In contrast to \sys's dynamic construction of
coallesced tasks at runtime, prior work has proposed to optimize task size
at compile time. CleanCut~\cite{cleancut_2018} program analysis statically
estimated energy consumption and split the program into tasks until all tasks
consumed less energy than the device can store.
%
An alternative program analysis~\cite{baghsorkhi_cgo_2018} generated different
versions of a program with different task sizes and empirically selected the
best among them~\cite{baghsorkhi_cgo_2018}.
%
HarvOS~\cite{mottola2017harvos} takes a hybrid approach that uses a program
analysis to place a minimal number of conditional checkpoints that test the
energy level at runtime before copying state, like Mementos~\cite{mementos}.

Unlike \sys, such compiler-based approaches, face the challenge of statically
predictiong energy consumption of arbitrary input-dependent code with
peripheral access, which is an problem without a general solution. Furthermore,
a static decomposition approach prevents portability across devices with
different storage capacitors. In contrast, \sys avoids forcing any assumptions
at compile time and adapts to energy storage capacity and incoming energy
conditions at runtime.

@@

A related domain is unbounded, page-based transactional memory and
deterministic parallel runtime systems~\cite{pagebasedtm,grace}.  These works
have a different mechanism and purpose than \sys---ensuring that data are
consistent and deterministically updated during concurrent executions.  Their
similarity with \sys is in managing state to ensure consistency at
the granularity of pages to amortize checking and tracking costs.
\sys's paging implementation, which keeps a shadow page for each page to use
during commit, is similar to the shadow paging scheme use for transactional
commit~\cite{pagebasedtm}.
